Malacology: A Programmable Storage System Built on Ceph

### 2016-05-10 paper/paper.tex
Removed section 2 and moved empirical evidence into intro

For example, a Ceph cluster contains a small number of
monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
system-wide metadata such as data distribution parameters and cluster
membership. The data distribution metadata maintained by
the cluster monitors is used by RADOS servers and clients to provide
consistent reads and writes, and both RADOS and the monitoring services
are used by a cluster of file system metadata servers that provide POSIX
semantics using sophisticated indexing and distributed locking services.
We contend that re-using and re-purposing these code-hardened subsystems
is paramount to successfully adapting storage systems to new APIs and
new storage devices without losing the benefits from years of
code-hardening work.

@@

First, we describe existing techniques for adapting storage systems
(\S\ref{problem}). Next, we discuss the ``distributed systems'' inspired
abstractions employed by Ceph (\S\ref{background}) and allude to their
re-usability as components in other services. The next section
enumerates the abstractions that Malacology exposes in the
implementation section (\S\ref{implementation}). We then show how we use
Malacology to synthesize entirely new storage services on an existing
system through configuration and small changes (\S\ref{services}). We
conclude by evaluating real-world use cases (\S\ref{evaluation}).

\section{Adapting Storage Systems}

\label{problem}

A consequence of complicated data management requirements and faster storage
devices is that storage systems may fail to meet the needs of applications.
Short of changing the appalication itself, approaches for meeting performance goals roughly fall into ``bolt-on'' services and changes to the storage system itself.

\subsection{Bolt-on services}
\label{bolt-on-services}

Bolt-on services are third-party solutions integrated into a
system to provide functionality and improve performance for common use-cases
(e.g. a metadata service). The introduction of additional services may provide
simplicity and performance enhancements, but ``bolt-on'' services often
duplicate functionality, unnecessarily increasing the likelihood of bugs or
introducing unpredictable performance problems. For example, we are developing
a high-performance distributed shared-log on Ceph called ZLog, based on the
CORFU protocol~\cite{balakrishnan_corfu_2012}. The CORFU protocol separates the concern of correctness from the concern of performance, enables serialization using a \emph{volatile}
``sequencer'' service. This high-performance in-memory counter is used to assign a log
position to each client that appends to the log. The sequencer is purely a performance optimization and does not have a role in correctness (if the sequencer fails, the tail of the log can always be retrieved by other means). Thus, it is the volatility of the sequencer that enables high performance (see Section~\ref{sec:zlog} for more detail). A ``bolt-on'' option for building
a distributed log is to use PAXOS, or an out-of-the-box solution such as
Zookeeper. However, these systems adhere to the preconceived notions of
storage, namely that data is safe from data loss, which is exactly the feature
that causes traditional approaches to run at significantly reduced speed.

\if{false}
\subsection{Application Changes}\label{application-changes}

The second approach that applications may use to adapt to a storage system
limitation is to extend the responsibility of the application. This means
changing the application itself by adding more data management intelligence or
using domain-specific middleware (e.g., a new data layout). For instance, an
application may be changed to exploit data locality or I/O parallelism in a
distributed storage system.
In SciHadoop~\cite{buck:hpc2011-scihadoop,buck:sc2013-scidr} both the Hadoop
application and the Hadoop framework itself are changed to leverage the structure of
scientific data (3D array-based data, more specifically) to
improve locality, performance, and increase the number of early results. %This is not a
%bad proposition, but 
However, changing applications creates a coupling that is highly tied to the
underlying physical properties of the system, making it difficult to
adapt to future changes at the storage system level.
\fi

\subsection{Storage System Changes}\label{storage-changes}

Storage system behaviour can be altered using two
techniques: tuning the system or introducing changes to the system.
%The
%difficulty of both of these approaches has given rise to a third
%technique called active storage.

\subsubsection{Tuning}\label{tuning}

Tuning large and complicated systems is difficult because of the number
of parameters (e.g., Hadoop 2.7.1 has 210 tunables) and because the
tunables are difficult to understand or quantify (e.g., Ceph tunables~\cite{sevilla:sc15-mantle}). To succesfully tune the system, a
developer must have domain and system-specific knowledge. Without this
intimate knowledge, the tuning turns into somewhat of a black art, where
the only way to figure out the best settings is trial and error.

Auto-tuning techniques attempt to find a good solution among a huge space of
available system configurations. However, in practice auto-tuning is limited
to only the configuration tunables that the storage system exposes and can be
overwhelmed with too many parameters.  For instance, auto-tuning may be
capable of identifying instances in which new data layouts would benefit a
workload, but unless the system can provide such a transformation, the option
is left off the table.  Starfish~\cite{herodotou:cidr2011-starfish} made
an attempt at this for Hadoop but this technique would need to severely limit
the space of parameters in order to be feasible, similar to the approach used in~\cite{behzad:sc2013-autotuning}.

\subsubsection{Software Changes}\label{software-changes}

To go beyond tuning one has to alter the storage system
itself. In the best case a developer is using an open-source system in which
case one must become familiar with often enormous code-bases before being able
to make changes, but then face the task of approaching a community to
integrate changes, or be forced to maintain an in-house version and risk
maintaining features that become incompatible.  A successful example of this
is the work that focused on HDFS scalability for metadata-intensive workloads~\cite{shvachko_hdfs_2010}. This has lead to modifications to its
architecture or API~\cite{balmin:sigmod2012-clydesdale} to improve
performance.

Alternatively, some systems have formal support for introducing additional
functionality, often in the form of a plugin infrastructure. Ceph provides the
ability to create domain-specific logical object interfaces by composing
existing interfaces. These interfaces are written in C++ and are statically
loaded into the system. %An example usage may be to atomically combine a
%``write'' with an index update to create a content-addressable object
%interface.

\if{false}
As an example use-case, ZLog needs to
store metadata information with each object, so for each operation, it
checks the epoch value and write capability (2 metadata reads) and
writes the index (metadata write).
\fi

For example, Figure~\ref{fig:datavsxattr} compares the throughput
(\emph{y} axis) over time (\emph{x} axis) of two OSD implementations of the same storage object interface for
storing metadata: either in the header of the byte stream (data) or in the
object extended attributes (XATTR). The speed for appending data without
any metadata operations (data raw) is also shown as a baseline for
comparison. For append-heavy workloads, storing metadata as a header in
the data performs about 1.5x better than storing metadata as an extended
attribute.

\begin{figure}[htbp]
\centering
\includegraphics{figures/cls_xattr-vs-data.png}
\caption{{[}\href{https://github.com/michaelsevilla/malacology-popper/blob/master/experiments/figure1/visualize.ipynb}{source}{]}
When appending data to objects, the object class that stores metadata in
a header in the data byte stream (data) performs 1.5x better than the
object class that stores metdata in the extended attributes of the
object (XATTR); it is almost as fast as appending data without updating
metadata (data raw).}
\label{fig:datavsxattr}
\end{figure}

The object interface facility provided by Ceph is more than
a proof-of-concept implementation resembling active storage.  Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

\if{false}
\begin{figure}[htbp]
\centering
\includegraphics{figures/obj-int-dev-churn.png}
\caption{Source code changes over time indicate the dynamic interface
development will need to support a high frequency of interface churn.
Each dot represents a Ceph commit with a corresponding number of lines
of code changed. \label{fig:obj-int-dev-churn}}
\end{figure}
\fi

\if{false}
These implementations use the active storage interfaces in Ceph. We
refer to this as \emph{storage programmability} which is a method by
which an application communicates its requirements to the storage system
in a way that allows the application to realize a new behavior without
sacrificing the correctness of the underlying system. This evaluation
and resulting implementation would have been a burden without object
classes.
\fi

@@

>>>>

\if{false}
For example, a Ceph cluster contains a small number of
monitoring nodes (MONs) that use PAXOS to maintain a consistent view of
system-wide metadata such as data distribution parameters and cluster
membership. The data distribution metadata maintained by
the cluster monitors is used by RADOS servers and clients to provide
consistent reads and writes, and both RADOS and the monitoring services
are used by a cluster of file system metadata servers that provide POSIX
semantics using sophisticated indexing and distributed locking services.
We contend that re-using and re-purposing these code-hardened subsystems
is paramount to successfully adapting storage systems to new APIs and
new storage devices without losing the benefits from years of
code-hardening work.
\fi

Ceph also provides empirical evidence that developers embrace programmable storage: Ceph already supports the
implementation of domain-specific logical object interfaces by composing
existing interfaces. These interfaces are written in C++ and are statically
loaded into the system. Figure~\ref{fig:obj-int-dev-growth} shows a dramatic growth in the production use of
domain-specific interfaces in the Ceph community since 2010.  What is most
remarkable is that this trend contradicts the notion that API changes are a
burden for users.
Rather it appears that a gap in existing interfaces are
being addressed through programmability. In fact, Table~\ref{table:objclasses}
categorizes existing interfaces and we clearly see a trend towards reusable
services.

@@

First, we describe existing techniques for adapting storage systems
(\S\ref{sec:evidence}). Next, we discuss the ``distributed systems'' inspired
abstractions employed by Ceph (\S\ref{background}) and allude to their
re-usability as components in other services. The next section
enumerates the abstractions that Malacology exposes in the
implementation section (\S\ref{implementation}). We then show how we use
Malacology to synthesize entirely new storage services on an existing
system through configuration and small changes (\S\ref{services}). We
conclude by evaluating real-world use cases (\S\ref{evaluation}).

@@

% IOStack: Software-Defined Object Storage
% sRoute: Treating the Storage Stack Like a Network
   % - lots of good refs to check out
% Information and control in gray-box systems
